---
title: "Report"
output: pdf_document
date: "2024-03-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#cat("\014")
#rm(list=ls())

# Pacman
if (!require("pacman")) install.packages("pacman"); library(pacman) 

# load data
load("/Users/hectorbahamonde/research/Inequality_Populism_Finland/dat.RData") # Load data
```


Let's take a look at the immigration patters in Finland. From the figure below, it is clear that immigration has risen.


```{r chunk:1}
p_load(car)
scatterplot(immigration.yearly ~ Year, 
            smooth=F, 
            regLine=F, 
            data=dat, 
            main="Yearly Immigrant Population in Finland", 
            xlab="Year", 
            ylab="N", 
            boxplots = F,
            legend=list(coords="bottomleft"),
            col="red")
```

And this is consistent with the literature: Populist far-right parties mobilize these grievances (i.e., higher immigration levels), gaining electoral support. However, what nobody seems to talk about is *"who"* are these immigrants, like, **where** do they come from... This is relevant as these grievances might be either *economic* or *cultural.* (1) Immigrants might be competing for the same (limited) jobs Finns perform, or maybe (2) people with different cultural backgrounds might ignite the so-called "cultural backlash." In any case, it is important to disentangle whether immigration might fuel either of these two grievances (economic, cultural) by analyzing the origin of immigrants. Concurrently, (1) do immigrants come from developed or underdeveloped countries? And also, (2) do immigrants come primarily from non-western "cultures"? Luckily, Finland has rich data on the origin of immigrant population.

# Immigration by Country of Origin

## Economic

The plot below shows that, in fact, immigration primarily comes from developed countries ("H"), and that the percentage of immigrants coming from poor or mid-poor countries ("L" and/or "LM") has systematically decreased. Thus, it might be safe to assume that Finland primarily "imports" educated immigrants. This might or might not help answering the "economic immigration grievance" hypothesis: skilled immigrants might be "taking over" Finnish jobs because of their high skills that might be transferable to the Finnish labor markets. In any case, the average immigrant is not necessarily of the "refugee" type, where public resources of the welfare state are necessarily spent to assist them (as the Finns claim). 

```{r chunk:2}
# load data
load("/Users/hectorbahamonde/research/Inequality_Populism_Finland/cat_econ_development_d.RData") # Load data

# plot
p_load(car)
par(pty="s")
scatterplot(freq~Year|Econ.Dev, smooth=F, regLine=T, data=cat.econ.development.d, 
            main="Immigration in Finland by Level\nof Economic Development of Country of Origin", 
            xlab="Year", 
            ylab="%",
            legend=list(coords="bottomleft"))
```


## Cultural

Now, one way to think about possible "cultural incompatibilities" is to look at the religion of immigrants. Coding the country of origin (Statistics Finland) with the primary religion of that country (World Bank datasets) helps us in doing so. The figure shows that the number of Muslim immigrants has grown quite a big deal. 

```{r chunk:3}
# load data
load("/Users/hectorbahamonde/research/Inequality_Populism_Finland/muslim_pop_imm_d.RData") # Load data

p_load(car)
options(scipen=999)
par(pty="s")

scatterplot(muslim.imm.yearly ~ Year, 
            smooth=F, 
            regLine=F, 
            data=muslim.pop.imm.d, 
            main="Yearly Muslim Population in Finland", 
            xlab="Year", 
            ylab="N", 
            boxplots = F,
            legend=list(coords="bottomleft"),
            col="red")
```


# Models: What's driving support for the far-right?

In the following tables I am estimating a bunch of models. All are multi-level models with city "fixed effects." Estimating models with city and time "fixed effects" give the same results. (For the paper we can have the without year FE in the Appendix). The DV is number of votes for the PS at the city level. The independent variables (in several combinations and groupings) are:

1. Gini: Yearly gini coefficient at the city level.

2. "HUM.imm": High and Upper-medium immigration. Yearly percentage of individuals who immigrate to Finland that come from High and Upper-medium economic development countries. 

3. Muslim immigration: Yearly immigration (count) classified by whether the immigrant's country of origin is predominantly Muslim.

4. Immigration yearly: Yearly total of immigration (count).

5. Others: 

- 1-year lag variables of each above.
- Interaction between "Gini" and "HUM.imm."


# Main Findings

1. Gini has the largest explanatory power no matter what it is controlled for.

2. Cultural hypothesis (i.e., Racism): Yes, total immigration increases PS votes, but only .1 points compared to *muslim* immigration, which is .4 points.

3. Economic explanation (1): Contrary to what the common mantra of "immigrants come to Finland to abuse the welfare state" or "take on the working class occupations," I not only find that immigrants in Finland come primary from developed countries (see figure), but also I find in the models that increasing levels immigration where the immigrant comes from a developed country ("HUM.imm")  significantly increases PS voting. This is a relatively specialized immigrant, not (necessarily) of the refugee type.

4. Economic explanation (2): Effects are just as large when inequality is interacted with skilled immigration (HUM.imm).

```{r chunk:4}
cat("\014")
rm(list=ls())
setwd("/Users/hectorbahamonde/research/Inequality_Populism_Finland/")

# Pacman
if (!require("pacman")) install.packages("pacman"); library(pacman) 

load("/Users/hectorbahamonde/research/Inequality_Populism_Finland/dat.RData")

# lme4
# https://rpubs.com/rslbliss/r_mlm_ws
p_load(lme4,Matrix, ggeffects, tidyverse,report)

options(scipen=999)



# 
m.1 <- lmer(PS ~ Gini +  HUM.imm + (1 | City), data = dat);# summary(m.1);ggpredict(m.1) %>% plot();report(m.1) # working hyp
m.2 <- lmer(PS ~ Gini +  muslim.imm.yearly + (1 | City), data = dat);# summary(m.2);ggpredict(m.2) %>% plot();report(m.2) # working hyp
m.3 <- lmer(PS ~ Gini +  immigration.yearly + (1 | City), data = dat);# summary(m.3);ggpredict(m.3) %>% plot();report(m.3) # working hyp
m.4 <- lmer(PS ~ Gini +  muslim.imm.yearly + immigration.yearly + (1 | City), data = dat);# summary(m.4);ggpredict(m.4) %>% plot();report(m.4) # working hyp
m.5 <- lmer(PS ~ Gini.lag.1 +  HUM.imm.lag.1 + (1 | City), data = dat);# summary(m.5);ggpredict(m.5) %>% plot();report(m.5) # working hyp
m.6 <- lmer(PS ~ Gini.lag.1 +  muslim.imm.yearly.lag.1 + (1 | City), data = dat);# summary(m.6);ggpredict(m.6) %>% plot();report(m.6) # working hyp
m.7 <- lmer(PS ~ Gini.lag.1 +  immigration.yearly.lag.1 + (1 | City), data = dat);# summary(m.7);ggpredict(m.7) %>% plot();report(m.7) # working hyp
m.8 <- lmer(PS ~ Gini.lag.1 +  muslim.imm.yearly.lag.1 + immigration.yearly.lag.1 + (1 | City), data = dat);# summary(m.8);ggpredict(m.8) %>% plot();report(m.8) # working hyp
m.9 <- lmer(PS ~ Gini *  HUM.imm + muslim.imm.yearly + (1 | City), data = dat);# summary(m.9);ggpredict(m.9) %>% plot();report(m.9) # working hyp
m.10 <- lmer(PS ~ Gini *  HUM.imm + immigration.yearly + (1 | City), data = dat);# summary(m.10);ggpredict(m.10) %>% plot();report(m.10) # working hyp



p_load(sjPlot,sjmisc,ggplot2)
plot_model(m.9, type = "int") + theme_bw() + theme(aspect.ratio=1)



# table
p_load(texreg)

screenreg( # use "screenreg" or "texreg" // BUT DO KEEP IT IN texreg FOR THE PAPER
  list(m.1, m.2, m.3, m.4)#, # list all the saved models here
  #omit.coef = "id"
)

screenreg( # use "screenreg" or "texreg" // BUT DO KEEP IT IN texreg FOR THE PAPER
  list(m.5, m.6, m.7,m.8)#, # list all the saved models here
  #omit.coef = "id"
)

screenreg( # use "screenreg" or "texreg" // BUT DO KEEP IT IN texreg FOR THE PAPER
  list(m.9,m.10)#, # list all the saved models here
  #omit.coef = "id"
)
```
